Multi-Modal Modular Sensorimotor Contingencies for Deformable Objects

### 2016-08-26 twostagedo.tex
changes to title etc

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Veronica E.~Arriola-Rios,
        Jeremy~L.~Wyatt}
%\author{Veronica E.~Arriola-Rios,~\IEEEmembership{Member,~IEEE,}
%        John~Doe,~\IEEEmembership{Fellow,~OSA,}
%        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
%\thanks{M. Shell is with the Department
%of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
%GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
%\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised December 27, 2012.}}

@@

% The paper headers
\markboth{IEEE CDS}%
{Arriola-Rios \MakeLowercase{\textit{et al.}}: Multi-Modal Modular Sensorimotor Contingencies for Deformable Objects}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{A} major challenge in robotic manipulation is to plan actions with objects so as to deform them into new shapes. Most objects that are manipulated by animals in natural environments are deformable, including objects that are elastic, plastic, breakable, tearable, and spreadable. \comment{The work reported here is based on the premise that in order to plan manipulations of deformable objects, the sensorimotor contingencies governing object deformation are required. A sensorimotor contingency is simply defined as the relationship between the robot's actions and the resultant changes in its observations. Thus different materials (elastic, plastic) present have different sensorimotor contingencies. The main hypothesis is that predictive models of these sensorimotor contingencies can be learned from data. These learned models can then be used to make the predictions required for planning. The predictions of a learned model can be fed back into the model, to predict an object's deformation behaviour over a variety of time scales. Without such predictions planners cannot reason about the effects of actions, or sequences of actions, to achieve desired effects.

Various models of deformation are used in engineering, such as finite element models, but these typically require careful optimisation by hand, and cannot run in real time. In graphics there are many methods used for generating qualitatively plausible simulations of deformable objects, but these are not easy to fit to real data. In contrast, robots must learn model parameters from data, on-line, in reasonable amounts of training time. This includes learning from the limited sensing typically available, such as partial visual views of an object, and force-torque monitoring only at contact points. Thus, tackling the prediction problem in robotics adds additional difficulties to an already challenging problem.}

@@

In summary the framework we present gives us the ability to solve {\em simultaneously} for the first time the following problems. Our model enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed. In the rest of the paper we present \comment{an overview of the model in more detail (Section~\ref{sec:model_overview});} related work (Section~\ref{sec:background}); describe the model (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); and the experiments on prediction and classification with both elastic and plastic objects (Section~\ref{sec:experiments}); finishing with a short discussion (Section~\ref{sec:discussion}).

@@

\begin{figure}[!t]
\centering
\setlength\fboxsep{0pt}    % padding
\setlength\fboxrule{0.5pt} % line width
\fbox{\includegraphics[width=3.5in]{arrio3.pdf}}
\caption{During training time, the system learns to predict the object's shape and reaction forces from real world data.}
\label{fig:system}
\end{figure}

@@

@@

\begin{table}[htp]
  %\renewcommand{\arraystretch}{1.3}
  \caption{Advances over previous work}\label{t:advances}
  \begin{tabular}{lccccccccc}
  %\hline\hline \\ [-1.5ex]
  \hline \\ [-1.5ex]
 & \begin{sideways}\textbf{Model type}\end{sideways} & \begin{sideways}\textbf{Learned from GT}\end{sideways} & \begin{sideways}\textbf{GT=Real data}\end{sideways} & \begin{sideways}\textbf{Prediction of Shape} \end{sideways} & \begin{sideways}\textbf{Prediction of Forces}\end{sideways} & \begin{sideways}\textbf{Actuator Retrieval} \end{sideways} & \begin{sideways}\textbf{Elastic}\end{sideways} & \begin{sideways}\textbf{Plastic}\end{sideways} & \begin{sideways}\textbf{Classification}\end{sideways} \\
 \hline\hline \\ [-1.5ex]
  Howard & MS & \gtick  & \gtick  & \rcross & \gtick & \rcross & \gtick & \rcross & \rcross\\
Teschner & MS & \rcross & \rcross & \gtick & \rcross & \gtick & \gtick & \gtick & \rcross \\
  Morris & MS & \gtick  & \rcross & \gtick & \rcross & \rcross & \gtick & \rcross & \rcross \\
  Frank & FEM & \gtick  & \gtick & \gtick & \rcross & QSA & \gtick & \rcross & \rcross \\
  This work & MS & \gtick  & \gtick & \gtick & \gtick & \gtick & \gtick & \gtick & \gtick \\
         & + &&&&&&&&\\
         & Regression &&&&&&&& \\
 %[-1.5ex] \\ \hline\hline \\ [-1.5ex]
 [-1.5ex] \\ \hline \\ [-1.5ex]
 \end{tabular}
 *MS = \textit{Mass-spring}\\
 *QSA = \textit{Quasi-Static Assumption}
\end{table}

Other interesting variants are: Conti et al.\ \cite{Conti2003} introduce six-degrees of freedom macroscopic elastic spheres described by mass, inertial and volumetric properties that are used to approximate the volume of deformable objects. The open source project Chai3D includes an implementation of this model. The spheres are placed along the skeleton of the object and are connected together with elastic links which model elongation, flexion and torsion properties. Each vertex of the surface mesh is attached to the nearest sphere or link with a damped spring.  In \cite{Burion2008}, Burion and Baur, collaborated with Conti's group to automatically calibrate this model using particle filters.  Again, there is no analysis of the dynamics and no real objects are used.  Furthermore, this model is more adequate for objects that deform preferably around fixed joints.

@@

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
% could we change the title to this? What do you think? I wanted to get the modular idea in there.
\title{Multi-Modal Modular Sensorimotor Contingencies for Deformable Objects
%A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation
}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Veronica E.~Arriola-Rios,
        Jeremy~L.~Wyatt \thanks{V.E. Arriola-Rios is with Facultad de Ciencias, UNAM, Mexico, e-mail: v.arriola@ciencias.unam.mx}
        \thanks{J.L. Wyatt is with the University of Birmingham, Edgbaston, Birmingham, B15 2TT, UK. e-mail: jeremy.l.wyatt@gmail.com }}%
% <-this % stops a space
%\thanks{J.L. Wyatt is with the University of Birmingham, Edgbaston, Birmingham, B15 2TT, UK. e-mail: jeremy.l.wyatt@gmail.com }% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised December 27, 2012.}}

@@

% The paper headers
\markboth{IEEE CDS}%
{Arriola-Rios and Wyatt: Multi-Modal Modular Sensorimotor Contingencies for Deformable Objects}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{A} major challenge in robot manipulation is to plan actions with objects so as to deform them into new shapes. Most objects that are manipulated by animals in natural environments are deformable, including objects that are elastic, plastic, breakable, tearable, and spreadable. \comment{The work reported here is based on the premise that in order to plan manipulations of deformable objects, the sensorimotor contingencies governing object deformation are required. A sensorimotor contingency is simply defined as the relationship between the robot's actions and the resultant changes in its observations. Different materials (elastic, plastic) have different sensorimotor contingencies. The main hypothesis is that predictive models of these sensorimotor contingencies can be learned from data. The predictions of a learned model can be fed back into the model, to predict an object's deformation behaviour over a variety of time scales. These learned models can then be used to make the predictions required for planning manipulation.}

\comment{There are various models of deformation used in engineering, such as finite element models, but these typically require careful optimisation by hand, and cannot run in real time. In graphics there are many methods used for generating qualitatively plausible simulations of deformable objects, but these are not easy to fit to real data. In contrast, robots must learn model parameters from data, on-line, in reasonable amounts of training time. This includes learning from the limited sensing typically available, such as partial visual views of an object, and force-torque monitoring only at contact points. Thus, tackling the prediction problem in robotics adds additional difficulties to an already challenging problem.}

@@

In summary, the framework we present gives us the ability to solve {\em simultaneously} for the first time the following problems. Our model enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed. In the rest of the paper we present related work (Section~\ref{sec:background}); \comment{give an overview of the framework (Section~\ref{sec:model_overview});} describe the two component models in detail (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); report experiments on prediction and classification with elastic and plastic objects (Section~\ref{sec:experiments}); and finish with a short discussion (Section~\ref{sec:discussion}).

@@

@@

\begin{table}[t]
  %\renewcommand{\arraystretch}{1.3}
  \caption{Advances over previous work}\label{t:advances}
  \begin{tabular}{lccccccccc}
  %\hline\hline \\ [-1.5ex]
  \hline \\ [-1.5ex]
 & \begin{sideways}\textbf{Model type}\end{sideways} & \begin{sideways}\textbf{Learned from GT}\end{sideways} & \begin{sideways}\textbf{GT=Real data}\end{sideways} & \begin{sideways}\textbf{Prediction of Shape} \end{sideways} & \begin{sideways}\textbf{Prediction of Forces}\end{sideways} & \begin{sideways}\textbf{Actuator Retrieval} \end{sideways} & \begin{sideways}\textbf{Elastic}\end{sideways} & \begin{sideways}\textbf{Plastic}\end{sideways} & \begin{sideways}\textbf{Classification}\end{sideways} \\
 \hline\hline \\ [-1.5ex]
  Howard & MS & \gtick  & \gtick  & \rcross & \gtick & \rcross & \gtick & \rcross & \rcross\\
Teschner & MS & \rcross & \rcross & \gtick & \rcross & \gtick & \gtick & \gtick & \rcross \\
  Morris & MS & \gtick  & \rcross & \gtick & \rcross & \rcross & \gtick & \rcross & \rcross \\
  Frank & FEM & \gtick  & \gtick & \gtick & \rcross & QSA & \gtick & \rcross & \rcross \\
  This work & MS & \gtick  & \gtick & \gtick & \gtick & \gtick & \gtick & \gtick & \gtick \\
         & + &&&&&&&&\\
         & Regression &&&&&&&& \\
 %[-1.5ex] \\ \hline\hline \\ [-1.5ex]
 [-1.5ex] \\ \hline \\ [-1.5ex]
 \end{tabular}
 *MS = \textit{Mass-spring}\\
 *QSA = \textit{Quasi-Static Assumption}
\end{table}

@@

Other interesting work includes that of Conti et al.\ \cite{Conti2003} which introduced six-degree of freedom macroscopic elastic spheres described by mass, inertial and volumetric properties that are used to approximate the volume of deformable objects. The open source project Chai3D includes an implementation of this model. The spheres are placed along the skeleton of the object and are connected together with elastic links which model elongation, flexion and torsion properties. Each vertex of the surface mesh is attached to the nearest sphere or link with a damped spring.  In \cite{Burion2008}, Burion and Baur, collaborated with Conti's group to automatically calibrate this model using particle filters.  Again, there is no analysis of the dynamics and no real objects are used.  Furthermore, this model is more adequate for objects that deform preferably around fixed joints.

@@

\begin{figure}[!t]
\centering
\setlength\fboxsep{0pt}    % padding
\setlength\fboxrule{0.5pt} % line width
\fbox{\includegraphics[width=3.5in]{modular.png}}
\caption{The modular architecture, in which each context /material specific module comprises two learned models of sensorimotor contingencies. The first predicts the force experienced for some motion and material. The second predicts the deformation of the object given the applied force. The output of the first predictor gives the input to the second predictor.}
\label{fig:system}
\end{figure}

\begin{figure}[t]
\centering
\setlength\fboxsep{0pt}    % padding
\setlength\fboxrule{0.5pt} % line width
\fbox{\includegraphics[width=3.5in]{arrio3.pdf}}
\caption{During training time, the system learns to predict the object's shape and reaction forces from real world data.}
\label{fig:system}
\end{figure}
